The \dealii Library, Version 9.1

### 2019-06-11 9.1/paper.tex
Some additional text improvements

For the 9.1 release, the geometric multigrid facilities have been extended and
revised for performance. The geometric multigrid algorithms for uniform and
adaptively refined meshes in \dealii are based on so-called local coarsening,
i.e., smoothening is applied level-per-level, skipping parts of the domain
where the mesh is not as refined. The algorithm for the assignment of the
owner on level cells and the implications on load balancing have been analyzed
in detail in \cite{ClevengerHeisterKanschatKronbichler2019}. While most of the
functionality was already available since the 8.5 release of \dealii presented
in \cite{dealII85}, several components have been finalized, such as the
support for certain renumbering algorithms that are beneficial for matrix-free
execution, and interfaces that allow the combination with matrix-free GPU
computations according to \cite{KronbichlerLjungkvist2019}.

A number of data structures and implementations in \dealii have been adapted
to ensure scalability of the matrix-free algorithms and geometric multigrid
infrastructure on more than 100,000 MPI ranks. Geometric multigrid solver for
the Poisson equation as described in \cite{KronbichlerWall2018} have been used
as performance tests during the acceptance phase of the SuperMUC-NG
supercomputer in Garching, Germany. Scaling tests have been performed on up to
the full machine with 304,128 cores of the Intel Xeon Skylake architecture and
an arithmetic performance of around 5 PFlop/s for a geometric multigrid solver
with polynomials of degree 4 has been reached. Compared to the official
LINPACK performance of the machine of 19.5 PFlop/s (the machine is listed on
position 8 of the top-500 list of November 2018), this can be considered an
extremely good value for a PDE solver which have classically only reached a
few percent of the LINPACK performance.
More importantly, this is achieved within a flexible framework supporting
arbitrary polynomial order on adaptively refined, unstructured meshes and
with algorithms in the matrix-free module of \dealii designed to minimize time
to solution and scalability, rather than maximizing the number of floating point operations. The
largest Poisson problem that has been solved on 304k cores contained 2.15
trillion unknowns (or 7.1 million unknowns per MPI rank) and was solved in 3.5
seconds. Also, CFD production runs with up to $10^{11}$ unknowns and $10^5$
time steps have been completed in less than seven hours, demonstrating the
capabilities of \dealii for large-scale parallel computations. The scaling tests
also revealed several relatively expensive operations in the setup of the
multigrid unknowns in \dealii's \texttt{DoFHandler} and \texttt{MGTransfer}
classes. While a few bottlenecks have already been resolved, we plan several
further improvements in setup times for the next release.

@@

\item The \texttt{MCMC-Laplace} code gallery program is a code useful
  for the forward solution used as a building block in
  Bayesian inverse problems, and for sampling the parameter space
  through a Metropolis-Hastings sampler (a kind of Monte Carlo
  Markov Chain method).
\end{itemize}

>>>>

For the 9.1 release, the geometric multigrid facilities have been extended and
revised for performance. The geometric multigrid algorithms for uniform and
adaptively refined meshes in \dealii are based on so-called local coarsening,
i.e., smoothening is done level-per-level, skipping parts of the domain
where the mesh is not as refined. The algorithm for the assignment of the
owner on level cells and the implications on load balancing have been analyzed
in detail in \cite{ClevengerHeisterKanschatKronbichler2019}. While most of the
functionality has already been available since the 8.5 release of \dealii presented
in \cite{dealII85}, several components have been finalized, such as the
support for certain renumbering algorithms that are beneficial for matrix-free
execution, and interfaces that allow the combination with matrix-free GPU
computations as showcased in \cite{KronbichlerLjungkvist2019}.

A number of data structures and implementations in \dealii have been adapted
to ensure scalability of the matrix-free algorithms and geometric multigrid
infrastructure on more than 100,000 MPI ranks. A geometric multigrid solver for
the Poisson equation as described in \cite{KronbichlerWall2018} has been used
as a performance test during the acceptance phase of the SuperMUC-NG
supercomputer in Garching, Germany. Scaling tests have been performed on up to
the full machine with 304,128 cores of the Intel Xeon Skylake architecture and
an arithmetic performance of around 5 PFlop/s for a geometric multigrid solver
with polynomials of degree 4 has been reached. Compared to the official
LINPACK performance of the machine of 19.5 PFlop/s (the machine is listed on
position 8 of the top-500 list of November 2018), this can be considered an
extremely good value for PDE solvers which have classically only reached a
few percent of the LINPACK performance.
More importantly, this is achieved within a flexible framework supporting
arbitrary polynomial order on adaptively refined, unstructured meshes and
with algorithms in the matrix-free module of \dealii designed to minimize time
to solution and scalability, rather than maximizing the number of floating point operations. The
largest Poisson problem that has been solved on 304k cores contained 2.15
trillion unknowns (or 7.1 million unknowns per MPI rank) and was solved in 3.5
seconds. Also, CFD production runs with $10^{11}$ unknowns and $10^5$
time steps have been completed in less than seven hours, demonstrating the
capabilities of \dealii for large-scale parallel computations. The scaling tests
also revealed several relatively expensive operations in the setup of the
multigrid unknowns in \dealii's \texttt{DoFHandler} and \texttt{MGTransfer}
classes. While a few bottlenecks have already been resolved for the present
release, we plan several further improvements of the setup stage for the next release.

@@

\item The \texttt{MCMC-Laplace} code gallery program is a code useful
  for the forward solution used as a building block in
  Bayesian inverse problems, and for sampling the parameter space
  through a Metropolis--Hastings sampler (a kind of Monte Carlo
  Markov Chain method).
\end{itemize}
